Arrows for Parallel Computation

### 2018-01-12 paper/content/utilityFunctions.tex
Phil's comment: backends
fixed paper matrix multiplication

instance Future BasicFuture a () where
    put _ = put'
    get _ = get'
\end{code}
\caption{|BasicFuture| type and its |Future| instance for the |Par| Monad and GpH Haskell backends.}
\label{fig:BasicFuture}
\end{figure}

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel Arrow running on two instances of |f +++ g| as described in the paper. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{|parEval2| definition.}
	\label{fig:parEval2}
\end{figure}
Furthermore, Fig.~\ref{fig:torus_example_rest} contains the omitted definitions of |prMMTr| (which calculates $AB^T$ for two matrices $A$ and $B$), |splitMatrix| (which splits the a matrix into chunks), and lastly |matAdd|, that calculates $A + B$ for two matrices $A$ and $B$.
\begin{figure}[h]
\begin{code}
prMMTr m1 m2 = [[sum (zipWith (*) row col) | col <- m2 ] | row <- m1]

@@

matAdd = chunksOf (dimX x) $ zipWith (+) (concat x) (concat y)
\end{code}
	\caption{|prMMTr|, |splitMatrix| and |matAdd| definition.}
	\label{fig:torus_example_rest}
\end{figure} %$

>>>>

instance Future BasicFuture a () where
    put _ = put'
    get _ = get'
\end{code}
\caption{|BasicFuture| type and its |Future| instance for the |Par| Monad and GpH.}
\label{fig:BasicFuture}
\end{figure}

@@

The |parEval2| skeleton is defined in Figure~\ref{fig:parEval2}. 
We start by transforming the |(a, c)| input into a two-element list |[Either a c]| by first tagging the two inputs with |Left| and |Right| and wrapping the right element in a singleton list with |return| so that we can combine them with |arr (uncurry (:))|. Next, we feed this list into a parallel Arrow running on two instances of |f +++ g| as described in the paper. After the calculation is finished, we convert the resulting |[Either b d]| into |([b], [d])| with |arr partitionEithers|. The two lists in this tuple contain only one element each by construction, so we can finally just convert the tuple to |(b, d)| in the last step.
\begin{figure}[h]
\begin{code}
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g =
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{code}
	\caption{|parEval2| definition.}
	\label{fig:parEval2}
\end{figure}
Furthermore, Fig.~\ref{fig:torus_example_rest} contains the omitted definitions of |prMM| (sequential matrix multiplication), |splitMatrix| (which splits the a matrix into chunks), |staggerHorizontally| and |staggerVertically| (to pre-rotate the matrices), and lastly |matAdd|, that calculates $A + B$ for two matrices $A$ and $B$.
\begin{figure}[h]
\begin{code}
prMM :: Matrix -> Matrix -> Matrix
prMM m1 m2 = prMMTr m1 (transpose m2)
  where
    prMMTr m1' m2' = [[sum (zipWith (*) row col) | col <- m2' ] | row <- m1']

@@

staggerHorizontally :: [[a]] -> [[a]]
staggerHorizontally matrix = zipWith leftRotate [0..] matrix

staggerVertically :: [[a]] -> [[a]]
staggerVertically matrix = transpose $ zipWith leftRotate [0..] (transpose matrix)

leftRotate :: Int -> [a] -> [a]
leftRotate i xs = xs2 ++ xs1 where
    (xs1,xs2) = splitAt i xs

matAdd = chunksOf (dimX x) $ zipWith (+) (concat x) (concat y)
\end{code}
	\caption{|prMMTr|, |splitMatrix|, |staggerHorizontally|, |staggerVertically| and |matAdd| definition.}
	\label{fig:torus_example_rest}
\end{figure} %$
